---
title: "Activity 2 - Day 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

## Load the necessary packages

Let's load necessary packages for this activity. 

```{r}
library(tidyverse)
library(tidymodels)
```

## Load the data 

Let's load the data.

```{r}
hfi <- data.frame(readr::read_csv("https://www.openintro.org/data/csv/hfi.csv"))
```

This data set contains 1458 rows and 123 columns. Each row of the data set represents year. 

```{r}
hfi_2016 <- hfi %>% filter(year == "2016") 
```

To display the relationship between the pf_score and pf_expression_control, I would use scatter plot because both variables are quantitative variables.

```{r}
# pf_score vs pf_expression_control
hfi_2016 %>% ggplot(aes(x = pf_score,
                        y = pf_expression_control)) + 
  geom_point(alpha=0.20)+
  labs(title = "Relationship between the pf score and pf expression control",
       x = "Personal Freedom Score",
       y = "Personal Freedom Expression Control",
       caption = "Data source: The OpenIntro site") + 
  geom_smooth(se = FALSE, method = "lm", size = 1) +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(text = element_text(face = "bold"))
```

From the above plot, the relationship looks linear. I think I would be comfortable using a linear model to predict the personal freedom score If I know a country’s pf_expression_control.

## Task 2: Overall model fit

### Sum of squared residuals
102.947 was the smallest sum of squares that I got with intercept of 4.1425 and x equals to 0.5709.


```{r}
m1 <- lm(pf_score ~ pf_expression_control, data = hfi_2016)

broom::tidy
tidy(m1)
```


### Interpretation: 

Interpret the intercept:

Intercept: For countries with a pf_expression_control of 0 (those with the largest amount of political pressure on media content), we expect their mean personal freedom score to be 4.28

Interpret the slope:

Slope: For every 1 unit increase in pf_expression_control, we expect a country's mean personal freedom score to increase 0.542 units.


```{r}
# Obtaining the correlation coefficient
cor(hfi_2016$pf_expression_control, hfi_2016$pf_score)
```

Since, the correlation coefficient is 0.85 we can say that there is a positive linear trend associated between pf_expression_control and personal freedom score. An increase in the value of one variable is associated with an increase in the other variable.

```{r}
# Assessing our model fit using R^2
broom::glance
glance(m1)
```

The value of R^2 for this model is 0.7141. This means 71.41% of the variability in the country’s personal freedom score is explained by their personal freedom expression control.

```{r}
# Fitting a new model that uses pf_expression_control to predict hf_score
m2 <- lm(hf_score ~ pf_expression_control, data = hfi_2016)

broom::tidy
tidy(m2)
```

From the above output, the equation of the regression line is:

yBar = 5.053396 + 0.368432*x

Where yBar = estimated hf_score and x = pf_expression_control

From the slope, it says that every 1 unit increase in pf_expression_control, we expect a country’s mean human freedom score to increase 0.37 units.

## Task 3: Prediction and prediction errors

```{r}
# pf_expression_control vs pf score
hfi_2016 %>% ggplot(aes(x = pf_expression_control,
                        y = pf_score)) + 
  geom_smooth(alpha=0.20) +
  labs(title = "Relationship between the  pf expression control and hf score",
       x = "Personal Freedom Expression Control",
       y = "Personal Freedom Score",
       caption = "Data source: The OpenIntro site") + 
  geom_smooth(se = FALSE, method = "lm", size = 1) +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(text = element_text(face = "bold"))
```

yBar = 4.2838153 + 0.5418452*3 = 5.909351

```{r}
# Check oberved values of pf_score with 6.7 rating for `pf_expression_control`

hfi_2016 %>%
  group_by(pf_score) %>%
  filter(pf_expression_control == 3)
```

residual = observed value - predicted value
observed value = I would consider the closest one with observed pf_score of less value for 3 rating for pf_expression_score.(Which we need to find from augment residual error)
predicted value = yBar = 5.909351
If the residual value is in negative then it is overestimated by that value.

## Task 4: Model diagnostics

```{r}
broom::augment

m1_aug <- augment(m1)
m1_aug
```

# Linearity:

```{r}
# Residuals vs. fitted (predicted) values
ggplot(data = m1_aug, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  xlab("Fitted values") +
  ylab("Residuals")
```

Linearity: From the above Residual by predicted value plot, there is no apparent pattern in the residuals plot. So it indicates there is a linear relationship between the two variables. Thus, the linearity assumption is met here.

# Nearly normal residuals: 

```{r}
# Histogram of the residuals.
ggplot(data = m1_aug, aes(x = .resid)) +
  geom_histogram(binwidth = 0.25) +
  xlab("Residuals")
```

```{r}
# QQ plot (Optional)
ggplot(data = m1_aug, aes(sample = .resid)) +
  stat_qq()
```

Nearly normal residuals: Based on the histogram, the residuals are nearly normal. Although slightly skewed to the left, there is only one real peak (uni-modal) on the bell curve in the residual histogram, so the nearly normal condition is met.

Constant variability: The points residuals vs. fitted plot show that points are scattered around 0, there is a constant variability.Thus, the constant variability condition appear to be met.
